---
title: "Final project"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(tidyverse)
library(caret)
library(ROCR)
library(MLmetrics)
library(mltools)
library(rpart.plot)
library(Hmisc)
library(corrplot)
```
importing datasets
```{r}
#https://www.kaggle.com/datasets/patelprashant/employee-attrition
attrition <- read.csv('/Users/peterlayne/Downloads/WA_Fn-UseC_-HR-Employee-Attrition 2.csv')
str(attrition)

```

```{r}
char <- names(select_if(attrition, is.character))
attrition[char] <- lapply(attrition[char], as.factor)
str(attrition)
table(attrition$JobRole)

intstofac <- c("Education", "EnvironmentSatisfaction", "JobInvolvement", "JobLevel", "JobSatisfaction", "PerformanceRating", "RelationshipSatisfaction", "StockOptionLevel", "WorkLifeBalance")
attrition[intstofac] <- lapply(attrition[intstofac], as.factor)
str(attrition)

attrition$PercentSalaryHike <- attrition$PercentSalaryHike/100
str(attrition)

#Dropping employee count, StandardHours as it is all the same values
attrition <- select(attrition,-c("EmployeeCount"))
attrition <- select(attrition,-c("StandardHours"))
```


#WHEN WE WANT TO NORMALIZE HERE IS THE CODE.
attrition_normalize <- preProcess(attrition, method=c("center","scale"))
norm1 <- predict(attrition_normalize, attrition)
summary(norm1)
str(norm1)



```{r}
#Generating corrplot

#generating df of only numeric values
numerics <- names(select_if(attrition, is.numeric))
attrition_numeric <- attrition[numerics]

corr_matrix_attrition <- cor(attrition_numeric)
corrplot(corr_matrix_attrition, method ="number")
#NEED THIS CORRPLOT TO SHOW CORRELATION OF VARIABLES TO ATTRITION WHICH IS A FACTOR SO ????

par(mar=c(1,1,1,1))

column_hists <- hist.data.frame(attrition_numeric, mtitl=TRUE)

  
  
#hist.data.frame(df) filter for numeric first
```

Normalize and select which to cluster with

attrition[numerics] <- lapply(attrition[numerics], normalize)
str(attrition)
attr_cluster <-attrition[numerics] 
str(attr_cluster )


```{r}
str(attrition)
attr <- caret::createDataPartition()
```

